Visual Diagnostics of a\kge{n} Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations

### 2020-12-30 paper.tex
removing extraneous text

\documentclass[AMS,STIX2COL]{WileyNJD-v2}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

% Set up for ASA data science journal
\articletype{Research Article}
\received{XX XX XXXX}
\revised{XX XX XXXX}
\accepted{XX XX XXXX}
\raggedbottom
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

% Paper header (title and authors)
\title{Visual Diagnostics of a\kge{n} \sout{Model} Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations}
\author[1]{Katherine Goode*}
\author[1,2]{Heike Hofmann}
\authormark{Goode and Hofmann}
\address[1]{\orgdiv{Department of Statistics}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\address[2]{\orgdiv{Center for Statistics and Applications in Forensic Evidence (CSAFE)}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\corres{*Katherine Goode, Department of Statistics, Iowa State University, Ames, IA. \email{kgoode@iastate.edu}}

% Abstract and keywords
\abstract[Abstract]{The importance of providing explanations for predictions made by black-box models has led to the development of \kge{explainer} model \sout{explainer} methods such as LIME (local interpretable model-agnostic explanations). LIME uses a surrogate model to explain the relationship between predictor variables and predictions from a black-box model in a local region around a prediction of interest. However, the quality of the resulting explanations relies on how well the explainer model captures the black-box model in a specified local region. Here we introduce three visual diagnostics to assess the quality of LIME explanations: (1) explanation scatterplots, (2) assessment metric plots, and (3) feature heatmaps. We apply the visual diagnostics to a forensics bullet matching dataset to show examples where LIME explanations depend on the tuning parameter values and the explainer model oversimplifies the black-box model. Our examples raise concerns about claims made of LIME that are similar to other criticisms in the literature.}
\keywords{explainable machine learning, black-box models, interpretability, statistical graphics, data science}

% Citation information
\jnlcitation{
\cname{\author{Goode K.}, \author{H. Hofmann}, }
\cyear{2020},
\ctitle{Visual Diagnostics of a\kge{n} \sout{Model} Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations},
\cjournal{Stat Anal Data Min: The ASA Data Sci Journal},
\cvol{volume, number and page}.}

@@

\begin{figure*}[!thp]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=6.5in]{figure/figure-01-1}

}

\end{knitrout}
\caption{\kge{Hypothetical scenarios depicting the effect of different weights on LIME (Left) An example of an explainer model with appropriate weights to provide a faithful approximation of the complex model. (Right) An example of an explainer model for the same prediction but with different weights that results in an explanation unfaithful to the complex model.} \sout{A conceptual depiction of a faithful LIME explainer model in the immediate neighborhood of a prediction of interest \kge{for a hypothetical black-box model trained using two features: Feature 1 (left) and Feature 2 (right). The explainer model is a ridge regression weighted by the inverse Gower distance to the prediction of interest raised to the power of \ensuremath{5\times 10^{-17}}.}} \sout{The predictions from a hypothetical black-box model are plotted against the standardized values of the two hypothetical predictor variables. The diamond shaped points represent the location of a prediction of interest. The size and opacity of the circular points indicate the weight assigned based on the distance to the prediction of interest computed using the inverse of the Gower distance metric raised to the power of \ensuremath{5\times 10^{-17}}. The black lines represent a weighted ridge regression model used as an explainer model that reasonably captures the relationship between the black-box predictions and the features in a local region around the prediction of interest. That is, the explainer is faithful to the complex model and produces a reasonable explanation that Feature 1 plays a more important role in the prediction of interest than Feature 2 since the magnitude of the slope associated with Feature 1 is larger.}}
\label{fig:figure-01}
\end{figure*}

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models \sout{are often used for the latter purpose. These models} have proven to perform well \kge{in problems with the latter objective} \sout{in a wide range of prediction problems}, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes it difficult to trust the model, especially in areas with high stakes decisions such as \kge{the} medical and forensics sciences. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

Throughout this paper, we distinguish between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor variable. In contrast, define {\it explanability} as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, molnar:2019}. Some are specific to one type of model (e.g. \citep{simonyan:2013} and \citep{urbanek:2008}), and others are model-agnostic (e.g. \citep{fisher:2018} and \citep{strumbelj:2014}). In this paper, we focus on \kge{the} \sout{one specific} model-agnostic method\sout{:} \kge{of} LIME \citep{ribeiro:2016}.

LIME (local interpretable model-agnostic explanations) \sout{is a method that} uses a surrogate model to relate predictor variables to black-box model predictions \sout{(i.e. a model explainer)} \citep{ribeiro:2016}. \sout{We distinguish between the terms of model explainer and explainer model: by \emph{model explainer} we denote the method for explaining a complex model using a surrogate model, while the \emph{explainer model}, or simply the \emph{explainer}, is the surrogate model.} While some \kge{explainer} model\kge{s} \sout{explainers} focus on understanding a model at the global level, LIME claims to provide explanations for individual predictions (local). Additionally, LIME is designed to work with any model (model-agnostic) and to produce easily understandable results (explanations) \citep{ribeiro:2016}. Conceptually, LIME fits a simple (interpretable) model, the explainer model, \kge{to approximate} \sout{meant to capture the behavior of} the \sout{(}complex\sout{) black-box} model in a local region around a prediction of interest. The simple model \kge{is interpreted to identify the} \sout{then provides interpretable estimates for} variables that most influence\sout{d} the \kge{complex model} prediction\sout{made by the complex model}.

\kge{While t}\sout{T}he concept of LIME is relatively simple\sout{: use an interpretable model to approximate a complex model in a local region. However}, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{pedersen:2020} \citep{ribeiro:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve \kge{reasonable} \sout{accurate} explanations, the tuning parameter values selected \kge{shoud} \sout{need to} be assessed.

\kge{To demonstrate the effect of a tuning parameter on LIME, we consider two different explainer models applied to the same prediction. Figure \ref{fig:figure-01} depicts two plots of the predictions from a hypothetical black-box model versus the feature used to train the model. The location of a prediction of interest is indicated by the diamonds. Globally, there is a clear non-linear relationship between the predictions and the feature, but the relationship could be approximated by a linear model in a local region around the prediction of interest. In both scenarios, a linear regression weighted by the proximity from an observation to the prediction of interest is used as the explainer model, but the proximities are computed differently as depicted by the varying sizes of the observations in the two plots.}

\kge{On the left, the distance between an observation and the prediction of interest is computed using the Gower distance metric \citep{gower:1971} and then raised to the power of \ensuremath{5\times 10^{-17}} to emphasize a very local region. The distances are subtracted from 1 to obtain a proximity. Here, the explainer model is faithful to the complex model since it captures the linear relationship between the black-box predictions in an immediate neighborhood of the prediction of interest. The slope \sout{of 0.634} explains that the black-box predictions increase as the feature increases in the local region around the prediction of interest.}

\kge{For the scenario on the right, the computation of the proximities is the same except the distances are raised to a power of 1 (the default exponent in the \emph{lime} R package \citep{pedersen:2020}). This causes the observations that are further away from the prediction of interest to be given larger weights than the scenario on the left. Here, the explainer model does not capture the linear trend near the prediction of interest, but instead, the explainer captures a more global trend. Without an assessment of this explainer model, the slope \sout{of -0.181} would incorrectly explain that the predictions decrease as the feature increases near the prediction of interest.} %\sout{This cannot be said of the explainer model in \autoref{fig:figure-02}. In addition, the magnitude of the slope associated with Feature 2  () is larger than that of Feature 1 (). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in the prediction of interest.}

%\kge{To demonstrate the effect of tuning parameters on LIME,we consider two different explainer models applied to the same prediction. Suppose a hypothetical black-box model is trained using one feature. Both plots in Figure1show the black-box model predictions versus the feature used to train the model.The diamond shaped point in the plots indicates the location of the prediction of interest to explain. A weight ridge regression with the black-box predictions as the response variable and standardized versions the features as predictor variables will be used as the explainer model in both scenarios, but the weights will vary. Note that the standardized features allow for direct comparisons of the model coefficients.}

%\kge{Figure \ref{fig:figure-01} depicts the first scenario where the weights are computed as the inverse proximity of the point to the prediction of interest computed using the Gower distance metric \citep{gower:1971} raised to the power of good_gower_power to emphasize a very local region. The point size represents the weight, and the black lines depict the ridge regression model given the observed prediction of interest value from the opposite feature. For Feature 1 (left), the explainer model captures the relationship in an immediate neighborhood around the prediction of interest with a slope of coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3), and for Feature 2 (right), the ridge regression has a small slope of coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3) that accurately reflects the absence of any global or local relationship between the black-box predictions and Feature 2. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.}

%\sout{\autoref{fig:figure-01} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two hypothetical model predictor variables. The diamond shaped points represent the location of the prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. In the example of \autoref{fig:figure-01}, an exponent of good_gower_power is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of two features in the data as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.}

%\sout{The plot on the left of \autoref{fig:figure-01} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of the prediction of interest. The explainer model captures the relationship in an immediate neighborhood around the prediction of interest with a slope of coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3). The plot on the right shows no global or local relationships between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set as the observed value of the prediction of interest, and it has an appropriately small slope of coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3). The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.}

%\sout{The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{ribeiro:2020} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameter values selected need to be assessed.}

%\kge{Figure \ref{fig:figure-02} depicts the second explainer model scenario. Again, the weights are computed using the Gower distance metric, but in this situation, the distances are raise to a power of bad_gower_power (the default exponent in the \emph{lime} R package).} \sout{We consider an example where the choice of exponent for the weights is crucial to the quality of the explanation. The plots in Figure \ref{fig:figure-02} show the same data as \autoref{fig:figure-01}, but the Gower distance metric exponent is decreased to bad_gower_power (the default exponent in the \emph{lime} R package).} This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:figure-01}, the explainer model captures the relationship between the black-box predictions in an immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:figure-02}. In addition, the magnitude of the slope associated with Feature 2  (coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()) is larger than that of Feature 1 (coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in the prediction of interest.

% \begin{figure*}[!thp]
% <<figure-02, out.width = '6.25in', fig.width = 7.95, fig.height = 3.25, warning = FALSE, eval = FALSE>>=
% 
% # Specify the figure size (for determining font size)
% f2_ow = 6.25
% f2_fw = 7.95
% f2_fs = fs * (f2_fw / f2_ow)
% 
% # Plot of good explainer model
% ggplot() +
%   facet_grid(. ~ feature, switch = "x") +
%   geom_point(
%     data = lime_data_gathered %>% 
%       mutate(
%         feature = fct_recode(
%           feature,
%           "Feature 1 (standardized)" = "Feature 1",
%           "Feature 2 (standardized)" = "Feature 2"
%           )
%         ),
%     mapping = aes(
%       x = feature_stnd_value,
%       y = prediction,
%       size = distance_value,
%       color = distance
%     ), 
%     alpha = 0.6
%   ) +
%   geom_abline(
%     data = coefs_data %>% 
%       filter(case == "bad") %>% 
%       mutate(
%         feature = fct_recode(
%           feature,
%           "Feature 1 (standardized)" = "Feature 1",
%           "Feature 2 (standardized)" = "Feature 2"
%           )
%         ),
%     mapping = aes(intercept = int, slope = slope),
%     size = 1
%   ) +
%   scale_color_manual(values = c("grey30", NA), guide = "none") +
%   geom_point(
%     data = prediction_of_interest_gathered %>% 
%       mutate(
%         feature = fct_recode(
%           feature,
%           "Feature 1 (standardized)" = "Feature 1",
%           "Feature 2 (standardized)" = "Feature 2"
%           )
%         ),
%     mapping = aes(x = feature_value,
%                   y = prediction,
%                   fill = color),
%     color = "black",
%     size = 5,
%     shape = 23,
%     alpha = 0.75
%   ) +
%   scale_fill_manual(values = "#FAAA72") +
%   geom_text(
%     data = data.frame(
%       feature_stnd_value = 1.2,
%       prediction = 1.6,
%       feature = c("Feature 1 (standardized)", "Feature 2 (standardized)"),
%       slope = c(
%         paste(
%           "Slope:",
%           coefs_data %>%
%             filter(case == "bad",
%                    feature == "Feature 1") %>%
%             pull(slope) %>%
%             round(3)
%         ),
%         paste(
%           "Slope:",
%           coefs_data %>%
%             filter(case == "bad",
%                    feature == "Feature 2") %>%
%             pull(slope) %>%
%             round(3)
%         )
%       )
%     ),
%     mapping = aes(x = feature_stnd_value,
%                   y = prediction,
%                   label = slope),
%     family = ff,
%     size = f2_fs / 3
%   ) +
%   labs(
%     x = "",
%     y = "Black-Box Prediction",
%     title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
%     subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
%     fill = "",
%     alpha = "Weight",
%     size = "Weight"
%   ) +
%   theme_linedraw(base_family = ff, base_size = f2_fs) +
%   theme(
%     panel.grid.major = element_blank(),

>>>>

\documentclass[AMS,STIX2COL]{WileyNJD-v2}

@@

% Set up for ASA data science journal
\articletype{Research Article}
\received{XX XX XXXX}
\revised{XX XX XXXX}
\accepted{XX XX XXXX}
\raggedbottom

\usepackage{Sweave}
\begin{document}
\input{paper-concordance}

% Paper header (title and authors)
\title{Visual Diagnostics of a\kge{n}  Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations}
\author[1]{Katherine Goode*}
\author[1,2]{Heike Hofmann}
\authormark{Goode and Hofmann}
\address[1]{\orgdiv{Department of Statistics}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\address[2]{\orgdiv{Center for Statistics and Applications in Forensic Evidence (CSAFE)}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\corres{*Katherine Goode, Department of Statistics, Iowa State University, Ames, IA. \email{kgoode@iastate.edu}}

% Abstract and keywords
\abstract[Abstract]{The importance of providing explanations for predictions made by black-box models has led to the development of \kge{explainer} model  methods such as LIME (local interpretable model-agnostic explanations). LIME uses a surrogate model to explain the relationship between predictor variables and predictions from a black-box model in a local region around a prediction of interest. However, the quality of the resulting explanations relies on how well the explainer model captures the black-box model in a specified local region. Here we introduce three visual diagnostics to assess the quality of LIME explanations: (1) explanation scatterplots, (2) assessment metric plots, and (3) feature heatmaps. We apply the visual diagnostics to a forensics bullet matching dataset to show examples where LIME explanations depend on the tuning parameter values and the explainer model oversimplifies the black-box model. Our examples raise concerns about claims made of LIME that are similar to other criticisms in the literature.}
\keywords{explainable machine learning, black-box models, interpretability, statistical graphics, data science}

% Citation information
\jnlcitation{
\cname{\author{Goode K.}, \author{H. Hofmann}, }
\cyear{2020},
\ctitle{Visual Diagnostics of a\kge{n}  Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations},
\cjournal{Stat Anal Data Min: The ASA Data Sci Journal},
\cvol{volume, number and page}.}

@@

\begin{Schunk}
\begin{Sinput}
> 
> # # Simulate example data
> # set.seed(20190624)
> # lime_data <-
> #   data.frame(
> #     feature1 = sort(runif(250, 0, 1)),
> #     feature2 = sample(x = 1:250, size = 250, replace = FALSE)) %>%
> #   mutate(
> #     feature1_stnd = (feature1 - mean(feature1)) / sd(feature1),
> #     feature2_stnd = (feature2 - mean(feature2)) / sd(feature2),
> #     prediction = if_else(feature1 >= 0 & feature1 < 0.1,
> #                          (0.3 * feature1) + rnorm(n(), 0, 0.01),
> #                  if_else(feature1 >= 0.1 & feature1 < 0.3,
> #                          rbeta(n(), 1, 0.5),
> #                  if_else(feature1 >= 0.3 & feature1 < 0.5,
> #                          sin(pi* feature1) + rnorm(n(), 0, 0.5),
> #                  if_else(feature1 >= 0.5 & feature1 < 0.8,
> #                          -(sin(pi* feature1) + rnorm(n(), 0, 0.1)) + 1,
> #                  if_else(feature1 >= 0.8 & feature1 < 0.9,
> #                          0.5 + runif(n(), -0.5, 0.5),
> #                          0.5 + rnorm(n(), 0, 0.3))))))) %>%
> #   bind_rows(
> #     data.frame(feature1 = rep(1, 30) + rnorm(5, 0, 0.05),
> #            feature2 = rep(245, 30) + rnorm(5, 0, 1),
> #            prediction = rep(0, 30) + rnorm(5, 0, 0.05)) %>%
> #   mutate(feature1_stnd = (feature1 - mean(feature1)) / sd(feature1),
> #          feature2_stnd = (feature2 - mean(feature2)) / sd(feature2)))
> # 
> # # Specify a prediction of interest
> # prediction_of_interest <-
> #   data.frame(feature1 = 0.07, feature2 = 200) %>%
> #   mutate(feature1_stnd = (feature1 - mean(lime_data$feature1)) /
> #            sd(lime_data$feature1),
> #          feature2_stnd = (feature2 - mean(lime_data$feature2)) /
> #            sd(lime_data$feature2),
> #          prediction = 0.05,
> #          color = factor("Prediction \nof Interest"))
> # 
> # # Specify the gower exponents
> # good_gower_power <- 50
> # bad_gower_power <- 1
> # 
> # # Compute the good distances between the prediction of interest
> # # and all other observations
> # lime_data$distance_good <-
> #   (1 - gower_dist(x = prediction_of_interest %>%
> #                     select(feature1_stnd, feature2_stnd),
> #                   y = lime_data %>%
> #                     select(feature1_stnd,
> #                            feature2_stnd)))^good_gower_power
> # 
> # # Compute the bad distances between the prediction of interest
> # # and all other observations
> # lime_data$distance_bad <-
> #   (1 - gower_dist(x = prediction_of_interest %>%
> #                     select(feature1_stnd, feature2_stnd),
> #                   y = lime_data %>%
> #                     select(feature1_stnd,
> #                            feature2_stnd)))^bad_gower_power
> # 
> # # Prepare the data for plotting
> # lime_data_gathered <- lime_data %>%
> #   gather(feature, feature_stnd_value,
> #          feature1_stnd:feature2_stnd) %>%
> #   gather(distance, distance_value, distance_good:distance_bad) %>%
> #   mutate(distance = fct_recode(distance,
> #                                "good" = "distance_good",
> #                                "bad" = "distance_bad"),
> #          feature = fct_recode(feature,
> #                               "Feature 1" = "feature1_stnd",
> #                               "Feature 2" = "feature2_stnd"))
> # 
> # # Prepare the prediction of interest data for plotting
> # prediction_of_interest_gathered <- prediction_of_interest %>%
> #   select(-feature1, -feature2) %>%
> #   gather(feature, feature_value, feature1_stnd, feature2_stnd) %>%
> #   mutate(feature = fct_recode(feature,
> #                               "Feature 1" = "feature1_stnd",
> #                               "Feature 2" = "feature2_stnd"))
> 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> 
> # # Fit the good interpretable explainer model
> # explainer_good <-
> #   glmnet(x = lime_data %>% select(feature1_stnd, feature2_stnd)
> #          %>% as.matrix(),
> #          y = lime_data$prediction,
> #          alpha = 0,
> #          lambda = 1,
> #          weights = lime_data$distance_good)
> # 
> # # Fit the bad interpretable explainer model
> # explainer_bad <-
> #   glmnet(x = lime_data %>% select(feature1_stnd, feature2_stnd)
> #          %>% as.matrix(),
> #          y = lime_data$prediction,
> #          alpha = 0,
> #          lambda = 1,
> #          weights = lime_data$distance_bad)
> # 
> # # Join the coefficients from the explainer model into a dataframe
> # coefs_data <- data.frame(case = c("good", "bad"),
> #                          b0 = c(coef(explainer_good)[1],
> #                                 coef(explainer_bad)[1]),
> #                          b1 = c(coef(explainer_good)[2],
> #                                 coef(explainer_bad)[2]),
> #                          b2 = c(coef(explainer_good)[3],
> #                                 coef(explainer_bad)[3])) %>%
> #   mutate(feature1 = b0 + b2*prediction_of_interest$feature2_stnd,
> #          feature2 = b0 + b1*prediction_of_interest$feature1_stnd) %>%
> #   gather(key = feature, value = int, feature1:feature2) %>%
> #   mutate(slope = c(coef(explainer_good)[2],
> #                    coef(explainer_good)[3],
> #                    coef(explainer_bad)[2],
> #                    coef(explainer_bad)[3]),
> #          feature = fct_recode(feature,
> #                               "Feature 1" = "feature1",
> #                               "Feature 2" = "feature2"))
> 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> # Specify the gower exponents
> good_gower_power <- 0.00000000000000005
> bad_gower_power <- 1
> # Specify a prediction of interest
> poi <-
+   data.frame(
+     poi = "Prediction \nof Interest",
+     feature = 0.1,
+     prediction = 0.05
+   )
> # Simulate example data
> set.seed(20190624)
> concept_data <-
+   data.frame(poi = "Training Data", 
+              feature = sort(c(runif(100, 0, 1), runif(250, 0, 7)))) %>%
+   mutate(prediction = sin(feature) + rnorm(350, 0, 0.25)) %>%
+   bind_rows(poi)
> # Compute the good distances between the prediction of interest
> # and all other observations
> concept_data$weights_good <-
+   1 - (gower_dist(
+     x = poi %>% select(feature),
+     y = concept_data %>% select(feature)
+   )) ^ good_gower_power
> # Compute the bad distances between the prediction of interest
> # and all other observations
> concept_data$weights_bad <-
+   1 - (gower_dist(
+     x = poi %>% select(feature),
+     y = concept_data %>% select(feature)
+   )) ^ bad_gower_power
> # Fit the good and bad interpretable explainer models
> explainer_good <-
+   lm(prediction ~ feature, data = concept_data, weights = weights_good)
> explainer_bad <-
+   lm(prediction ~ feature, data = concept_data, weights = weights_bad)
> # Create labels for the gower powers
> good_power_label <-
+   paste0("Faithful Local Explainer \n(exponent used to compute weights: ",
+          good_gower_power,
+          ")")
> bad_power_label <-
+   paste0("Unfaithful Local Explainer \n(exponent used to compute weights: ",
+          bad_gower_power,
+          ")")
> # Join the coefficients from the explainer model into a dataframe
> coefs_data <-
+   data.frame(
+     case = c(good_power_label, bad_power_label),
+     b0 = c(coef(explainer_good)[1],
+            coef(explainer_bad)[1]),
+     b1 = c(coef(explainer_good)[2],
+            coef(explainer_bad)[2]),
+     feature = 6,
+     prediction = 1.6
+   ) %>%
+   mutate(text = paste0("Slope: ", round(b1, 3)))
> # Prepare data for plotting
> concept_data_long <-
+   concept_data %>%
+   pivot_longer(names_to = "case",
+                values_to = "weight",
+                cols = weights_good:weights_bad) %>%
+   mutate(case = ifelse(case == "weights_good", good_power_label, bad_power_label))
\end{Sinput}
\end{Schunk}

\begin{figure*}[!thp]
\begin{Schunk}
\begin{Sinput}
> # Specify the figure size (for determining font size)
> f1_ow = 6.5
> f1_fw = 7.95
> f1_fs = fs * (f1_fw / f1_ow)
> # Plot of good explainer model
> ggplot() +
+   facet_grid(. ~ case) +
+   geom_point(
+     data = concept_data_long,
+     mapping = aes(
+       x = feature,
+       y = prediction,
+       fill = poi,
+       color = poi,
+       shape = poi,
+       alpha = poi,
+       size = weight
+     )
+   ) +
+   geom_abline(
+     data = coefs_data,
+     mapping = aes(intercept = b0, slope = b1),
+     size = 1
+   ) +
+   scale_fill_manual(values = c("#FAAA72", "grey30")) +
+   scale_color_manual(values = c("grey30", "grey30")) +
+   scale_alpha_manual(values = c(0.75, 0.6)) +
+   scale_shape_manual(values = c(23, 21)) +
+   geom_text(
+     data = coefs_data,
+     mapping = aes(x = feature,
+                   y = prediction,
+                   label = text),
+     family = ff,
+     size = f1_fs / 2.85
+   ) +
+   labs(x = "Black-Box Model Feature",
+        y = "Black-Box Prediction",
+        size = "Weight", 
+        fill = "", 
+        color = "",
+        alpha = "",
+        shape = "") +
+   theme_linedraw(base_family = ff, base_size = f1_fs) +
+   theme(
+     panel.grid.major = element_blank(),
+     panel.grid.minor = element_blank(),
+     axis.text.x = element_blank(),
+     axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(),
+     axis.ticks.y = element_blank(),
+     strip.placement = "outside",
+     strip.background = element_rect(color = "white", fill = "white"),
+     strip.text.x = element_text(size = f1_fs, color = "black"),
+     plot.title = element_text(size = f1_fs)
+   ) +
+   guides(
+     fill = guide_legend(order = 1, override.aes = list(size = 5)),
+     color = guide_legend(order = 1),
+     shape = guide_legend(order = 1),
+     alpha = guide_legend(order = 1),
+     size = guide_legend(
+       order = 2,
+       override.aes = list(color = "grey30", alpha = 0.6)
+     )
+   )
> 
> # # Plot of good explainer model
> # ggplot() +
> #   facet_grid(. ~ feature, switch = "x") +
> #   geom_point(
> #     data = lime_data_gathered %>% mutate(
> #       feature = fct_recode(
> #         feature,
> #         "Feature 1 (standardized)" = "Feature 1",
> #         "Feature 2 (standardized)" = "Feature 2"
> #       )
> #     ),
> #     mapping = aes(
> #       x = feature_stnd_value,
> #       y = prediction,
> #       size = distance_value,
> #       color = distance
> #     ),
> #     alpha = 0.6
> #   ) +
> #   geom_abline(
> #     data = coefs_data %>%
> #       filter(case == "good") %>%
> #       mutate(
> #         feature = fct_recode(
> #           feature,
> #           "Feature 1 (standardized)" = "Feature 1",
> #           "Feature 2 (standardized)" = "Feature 2"
> #           )
> #         ),
> #     mapping = aes(intercept = int, slope = slope),
> #     size = 1
> #   ) +
> #   scale_color_manual(values = c(NA, "grey30"), guide = "none") +
> #   geom_point(
> #     data = prediction_of_interest_gathered %>% mutate(
> #       feature = fct_recode(
> #         feature,
> #         "Feature 1 (standardized)" = "Feature 1",
> #         "Feature 2 (standardized)" = "Feature 2"
> #       )
> #     ),
> #     mapping = aes(x = feature_value,
> #                   y = prediction,
> #                   fill = color),
> #     color = "black",
> #     size = 5,
> #     shape = 23,
> #     alpha = 0.75
> #   ) +
> #   scale_fill_manual(values = "#FAAA72") +
> #   geom_text(
> #     data = data.frame(
> #       feature_stnd_value = 1.2,
> #       prediction = 1.6,
> #       feature = c("Feature 1 (standardized)", "Feature 2 (standardized)"),
> #       slope = c(
> #         paste(
> #           "Slope:",
> #           coefs_data %>%
> #             filter(case == "good",
> #                    feature == "Feature 1") %>%
> #             pull(slope) %>%
> #             round(3)
> #         ),
> #         paste(
> #           "Slope:",
> #           coefs_data %>%
> #             filter(case == "good",
> #                    feature == "Feature 2") %>%
> #             pull(slope) %>%
> #             round(3)
> #         )
> #       )
> #     ),
> #     mapping = aes(x = feature_stnd_value,
> #                   y = prediction,
> #                   label = slope),
> #     family = ff,
> #     size = f1_fs / 2.85
> #   ) +
> #   labs(
> #     x = "",
> #     y = "Black-Box Prediction",
> #     title = "Conceptual Depiction of a Faithful Local Explainer Model",
> #     subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
> #     fill = "",
> #     alpha = "Weight",
> #     size = "Weight"
> #   ) +
> #   theme_linedraw(base_family = ff, base_size = f1_fs) +
> #   theme(
> #     panel.grid.major = element_blank(),
> #     panel.grid.minor = element_blank(),
> #     axis.text.x = element_blank(),
> #     axis.ticks.x = element_blank(),
> #     axis.text.y = element_blank(),
> #     axis.ticks.y = element_blank(),
> #     strip.placement = "outside",
> #     strip.background = element_rect(color = "white", fill = "white"),
> #     strip.text.x = element_text(size = f1_fs, color = "black"),
> #     plot.title = element_text(size = f1_fs)
> #   ) +
> #   guides(
> #     fill = guide_legend(order = 1),
> #     size = guide_legend(order = 2),
> #     alpha = guide_legend(order = 2)
> #   )
> 
\end{Sinput}
\end{Schunk}
